---
title: "p8105_hw3_rl3616"
author: "Ruipeng Li"
date: "`r format(Sys.Date())`"
output: github_document
---

# Problem 1

Loading data
```{r}
library(p8105.datasets)
library(dplyr)
library(ggplot2)
library(tidyr)
data("instacart") 
```

#### Part a)
The Instacart dataset contains 1,384,617 observations and 15 variables, representing individual items from customers’ grocery orders.

Each observation corresponds to one product within an order, with variables describing the order time `order_hour_of_day`, date in week `order_dow`, customer `user_id`, product details `product_name, aisle, department`, and purchase behavior `add_to_cart_order, reordered, days_since_prior_order`.

The dataset allows exploration of shopping habits such as the most popular aisles, frequently reordered products, and peak shopping hours.

#### Part b)
Also, there are `r n_distinct(instacart$aisle)` different aisles in the dataset, and 
```{r}
top_aisles <- instacart |> 
  count(aisle, sort = TRUE)
```

With this data, we can see the top 1 aisle that most items ordered from was "`r top_aisles$aisle[1]`" and follow up with "`r top_aisles$aisle[2]`" and "`r top_aisles$aisle[3]`"

#### Part c)
```{r}
aisle_plot <- top_aisles |> 
  filter(n > 10000) |> 
  ggplot(aes(x = reorder(aisle, n), y = n)) +
  geom_col(width = 0.7, fill = "steelblue") +
  geom_text(aes(label = scales::comma(n)), 
            hjust = -0.1, size = 2.5) +
  coord_flip() +
  labs(
    title = "Aisles with more than 10,000 items ordered",
       x = "Aisle", y = "Number of items ordered"
    ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal(base_size = 8)

ggsave("figures/Problem 1/aisle_plot.png", aisle_plot, width = 10, dpi = 300)
```

#### Part d)
```{r}
top_item_table <- instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |>
  summarise(times_ordered = n()) |> 
  slice_max(times_ordered, n = 3)
```

#### Part e)
```{r}
mean_hour_table <- instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarise(mean_hour = mean(order_hour_of_day)) |> 
  mutate(order_dow = factor(order_dow,
    levels = 0:6,
    labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
  )) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

