---
title: "p8105_hw3_rl3616"
author: "Ruipeng Li"
date: "`r format(Sys.Date())`"
output: github_document
---

# Problem 1

Loading data
```{r}
library(p8105.datasets)
library(ggplot2)
library(tidyverse)
data("instacart") 
```

#### Part a)
The Instacart dataset contains 1,384,617 observations and 15 variables, representing individual items from customers’ grocery orders.

Each observation corresponds to one product within an order, with variables describing the order time `order_hour_of_day`, date in week `order_dow`, customer `user_id`, product details `product_name, aisle, department`, and purchase behavior `add_to_cart_order, reordered, days_since_prior_order`.

The dataset allows exploration of shopping habits such as the most popular aisles, frequently reordered products, and peak shopping hours.

#### Part b)
Also, there are `r n_distinct(instacart$aisle)` different aisles in the dataset, and 
```{r}
top_aisles <- instacart |> 
  count(aisle, sort = TRUE)
```

With this data, we can see the top 1 aisle that most items ordered from was "`r top_aisles$aisle[1]`" and follow up with "`r top_aisles$aisle[2]`" and "`r top_aisles$aisle[3]`"

#### Part c)
```{r}
aisle_plot <- top_aisles |> 
  filter(n > 10000) |> 
  ggplot(aes(x = reorder(aisle, n), y = n)) +
  geom_col(width = 0.7, fill = "steelblue") +
  geom_text(aes(label = scales::comma(n)), 
            hjust = -0.1, size = 2.5) +
  coord_flip() +
  labs(
    title = "Aisles with more than 10,000 items ordered",
       x = "Aisle", y = "Number of items ordered"
    ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal(base_size = 8)

ggsave("figures/Problem 1/aisle_plot.png", aisle_plot, width = 10, dpi = 300)
```

#### Part d)
```{r}
top_item_table <- instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |>
  summarise(times_ordered = n()) |> 
  slice_max(times_ordered, n = 3)
```

#### Part e)
```{r}
mean_hour_table <- instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarise(mean_hour = mean(order_hour_of_day)) |> 
  mutate(order_dow = factor(order_dow,
    levels = 0:6,
    labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
  )) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

# Problem 2

#### Cleaning Zillow datasets(from homework 2)
```{r}

zip_zori_df <- read_csv("data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") |> 
  janitor::clean_names() |>
  rename(zip_code = region_name) |> 
  select(-region_type, -state_name, -state, -city, -metro) |> 
  pivot_longer(
    cols = starts_with("x20"),
    names_to = "date",
    values_to = "rent_price"
  ) |> 
  filter(!is.na(rent_price)) |>
  mutate(
    county_name = str_replace(county_name, " County", ""),
    date = str_replace(date, "x", ""),
    date = ymd(date),
    borough = case_when(
      county_name == "New York" ~ "Manhattan",
      county_name == "Kings" ~ "Brooklyn",
      county_name == "Queens" ~ "Queens",
      county_name == "Bronx" ~ "Bronx",
      county_name == "Richmond" ~ "Staten Island")
    ) |> 
   relocate("county_name","borough")

```

```{r}

zip_count <- zip_zori_df |> 
  group_by(zip_code, county_name) |> 
  summarise(n_observed = n()) |> 
  arrange(desc(n_observed))

```
Using this data, we can see there are `r sum(zip_count$n_observed == 116)` ZIP codes are observed 116 times, and there are `r sum(zip_count$n_observed < 10)` ZIP codes observed fewer than 10 times. 

Using arranged data we can see almost all ZIPs in Manhattan and Brooklyn appear 116 times;
while many in Staten Island or the Bronx might appear only a dozen or fewer times. I think it's really depends on population density. Which frequent appearances indicate rich data and active listings; few appearances indicate an unpopular area, a change in listings, or missing data.

#### Average rental price
```{r}
mean_price <- zip_zori_df |> 
  mutate(year = lubridate::year(date)) |>
  group_by(year, borough) |> 
  summarise(average_price = mean(rent_price)) |> 
  pivot_wider(
    names_from = year,
    values_from = average_price
  )
```
